---
title: "Part 3:	Iris Flower Data Analysis"
output: html_notebook
---

## Data analysis to be conducted in R.

*The	data set used is built-in the studio R package. It is called the "Edgar Anderson's	Iris	Data"	set.	The	data is in 'data frame' format and	named	'iris'.	Type 'iris'	in R Help	to find	out	more about the data	set.*

<br/>

### 1) Quote the description of the data set as found default R Help. Present the data briefly by indicating the variables names, their statistical class, the size of the vectors (1D arrays), and by showing first 6 lines of values for each variables.
<br/>

**This famous (Fisher's or Anderson's) iris data set gives the measurements in centimeters of the variables sepal length and width and petal length and width, respectively, for 50 flowers from each of 3 species of iris. The species are Iris setosa, versicolor, and virginica.**

```{r}
str(iris)
```
**The data is composed of 5 different variables. These variables are:**

* Sepal.Lenght which contains continuous numeric values
* Sepal.Width which contains continuous numeric values
* Petal.Length which contains continuous numeric values
* Petal.Width which contains continuous numeric values
* Species which contains categorical values
 
```{r}
total_records <- nrow(iris)
```

**There are a total of `r total_records` records. We can see the first six:**

```{r}
head(iris, 6)
```

<br/>

### 2) Calculate and present the standard descriptive statistics (central and dispersion) for all 4 numerical variables.

```{r}
num.vars <- iris[,1:4]
summary(num.vars)
```
- **Sepal length** spans from a minimum of 4.3cm to 7.9cm. The mean value is 5.843cm while the median is 5.8cm. The mean and median values are very close to each other, this could imply that the data is symmetrical. 
```{r}
boxplot(iris$Sepal.Length, horizontal=TRUE)
```
In the boxplot, we notice that the data is right skewed. We can calculate the standard deviation to see the dispersion of the feature.
```{r}
sd_s_l <- sd(iris$Sepal.Length)
```
**The standard deviation of Sepal Length is $\sigma=`r sd_s_l`$**.

Sepal length seems to exhibit a low variation.
<br/>
<br/>
- **Sepal width** as well as sepal length, has a median value which is very close to the mean. The 1st and 3rd quartiles seem to be equidistant.

```{r}
boxplot(iris$Sepal.Width, horizontal=TRUE)
```
Sepal width appears to be normally distributed and also shows few outliers.
```{r}
sd_s_w <- sd(iris$Sepal.Width)
```
**The standard deviation of Sepal Width is $\sigma=`r sd_s_w`$**.

<br/>
<br/>
- **Petal length** seems to be less symmetrical as the mean and median values differ: the median is significantly higher than the mean. Altough, we can also notice that the 1st quartile has a small value compared to the other quartiles meaning that there are many data points included in the small range.
```{r}
boxplot(iris$Petal.Length, horizontal=TRUE)
```
In fact, we can see that the data is right skewed.
```{r}
sd_p_l <- sd(iris$Petal.Length)
```
**The standard deviation of Petal Length is $\sigma=`r sd_p_l`$**.

It also has a higher dispersion than the two features previously examined.
<br />
<br />
- **Petal width** has similar central statistics to petal length, and as well shows many data points in the small range of values.
```{r}
boxplot(iris$Petal.Width, horizontal=TRUE)
```
In fact the distribution looks similar to the one of petal length.
```{r}
sd_p_w <- sd(iris$Petal.Width)
```
**The standard deviation of Petal Width is $\sigma=`r sd_p_w`$**.

In this case the standard deviation has a smaller value but it is to be expected taken the scale into consideration.

### 3) Consider all possible scatter plots of the 4 numerical variables against one another.

#### i) How many plots are possible in theory? How many scatter plots are there, if one uses each variable pair once? Use factorial based formulas.

We have a total of four variables and we need to pick two at the time. There is no repetition as it is not necessary to plot one variable against itself. 

$$^{4}C_{2}=\frac{n!}{k!(n-k)!}$$
```{r}
  n <- 4
  k <- 2
  max_plots <- factorial(n)/(factorial(k)*factorial(n-k))
```

**`r max_plots` is the total number of combinations for choosing `r k` elements at a time from `r n` distinct elements.**


#### ii) Plot all scatter plots using each variable pair once.

```{r}
pairs(num.vars,     
      pch = 20,  
      col = c("#04ADC5", "#BBD931", "#FB3F16")[iris$Species] ,
      upper.panel=NULL)
```


#### iii) Indicate, which plots seem to exhibit a linear relationship (straight line)? Conclude in one sentence what these relationships indicate.

1. Petal length vs. Petal width
2. Sepal length vs. Petal length
3. Sepal lenght vs. Petal width

<br/>

<span style="color: red;">Explanation needed</span>

### 4)

#### i) Create (relative) frequency histograms for each variable.

```{r}
hist(iris$Sepal.Width, freq=TRUE, main='Histogram of sepal width', xlab='Sepal width')
```

```{r}
hist(iris$Sepal.Length, freq=TRUE, main='Histogram of sepal length', xlab='Sepal length')
```

```{r}
hist(iris$Petal.Length, freq=TRUE, main='Histogram of petal length', xlab='Petal length')
```

```{r}
hist(iris$Petal.Width, freq=TRUE, main='Histogram of Petal width', xlab='Petal width')
```


#### ii) Which variable appears to be normally distributed? Below this will be variable X.

Sepal width

#### iii) Isolate the variable which appears normally distributed, X. Based on its mean and its standard deviation, define (in a formula) the model normal probability density function and plot it.

```{r}
X <- iris$Sepal.Width

X.mean <- mean(X)
```

**The mean is $\mu=`r X.mean`$**

```{r}
X.sd <- sd(X)
```

**The standard deviation is $\sigma=`r X.sd`$**

 $$f(x)=\frac{1}{\sigma\sqrt{2\pi}}e^{-\frac{(x-\mu)^{2}}{2\sigma^{2}}}$$

In R we can use the `dnorm` function.

```{r}
x.axis <- seq(1, 5, length=150)
y.axis <- dnorm(x.axis, mean=X.mean, sd=X.sd)
plot(x.axis, y.axis, type='l', main="Normal distribution",
  xlab="X", ylab="X ~ N(3.06, 0.44)")
```

<br/>

### 5) Consider the data in X as a population and assume it is normally distributed. Estimate the probability that iris flowers have:

<br/>

***A helper function to represent data as percentage:***

```{r}
percent <- function(number){
  return(sprintf("%1.2f%%", number * 100))
}
```
<br/>

#### i) X > 3
```{r}
p_i <- sum(X > 3)/length(X)
```
The probability is **`r p_i`** -> **`r percent(p_i)`**.

```{r}
pn_i <- pnorm(3, mean=X.mean, sd=X.sd, lower.tail=FALSE)
```
The probability of normal distribution is **`r pn_i`** -> **`r percent(pn_i)`**.

#### ii) 2.5 < X < 3.5
```{r}
p_ii <- sum(2.5 < X & X < 3.5)/length(X)
```
The probability is **`r p_ii`** -> **`r percent(p_ii)`**.

```{r}
pn_ii <- (pnorm(3.5, mean=X.mean, sd=X.sd, lower.tail=TRUE) - 
          pnorm(2.5,mean=X.mean, sd=X.sd, lower.tail=TRUE))
```
The probability of normal distribution is **`r pn_ii`** -> **`r percent(pn_ii)`**.

#### iii) X < 2.5
```{r}
p_iii <- sum(X < 2.5)/length(X)
```
The probability is **`r p_iii`** -> **`r percent(p_iii)`**.

```{r}
pn_iii <- pnorm(2.5, mean=X.mean, sd=X.sd, lower.tail=TRUE)
```
The probability of normal distribution is **`r pn_iii`** -> **`r percent(pn_iii)`**.

<br/>

### 6) Take the scatter plots of the variable X versus i) the petal length and ii) the sepal length. Notice in each plot the data points appear to be clustered in two clusters. Put a threshold (limit) to isolate the data points of the first cluster corresponding to lower values of petal length and petal width. Examine if these belong to same species, by checking the data variable 'Species'.

## *with X (sepal width)*
```{r}
plot(X, iris$Petal.Length, pch=19)
```

```{r}
plot(X, iris$Sepal.Length, pch=19)
```

## *with petal width*
```{r}
plot(iris$Petal.Width, iris$Petal.Length, pch=19)
```

```{r}
plot(iris$Petal.Width, iris$Sepal.Length, pch=19)
```

#### 2. Put a threshold (limit) to isolate the data points of the first cluster corresponding to lower values of petal length and petal width. Examine if these belong to same species, by checking the data variable 'Species'.

```{r}
indices <- iris$Petal.Width < 0.8 & iris$Petal.Length < 2.5
species <- unique(iris$Species[indices])
```
They all seem to belong to the same species - **`r species`**.

### 7)

#### i) Establish the 95% confidence interval for the true mean of X based on all the X data.

```{r}
margin_error <- qnorm(.975)*(X.sd/sqrt(length(X)))
ll_i <- X.mean - margin_error
ul_i <- X.mean + margin_error
```

The **lower limit is *`r ll_i`* ** and the **upper limit is *`r ul_i`* **

#### ii) Take a subset, S, of X that includes only the X values of the species that was identified in 3.6. Establish the 95% confidence interval for the true mean X based on this sample subset.

```{r}
S <- X[indices]
S.mean <- mean(S)
S.sd <- sd(S)
margin_error <- qnorm(.975)*(S.sd/sqrt(length(S)))
ll_ii <- S.mean - margin_error
ul_ii <- S.mean + margin_error
```
The **lower limit is *`r ll_ii`* ** and the **upper limit is *`r ul_ii`* **

#### iii) Are both samples large enough for such statistical study? What type of statistical study is this, what does it consist in, and how does it compare to the approach used in 3.5?

<span style="color: red;">Explanation needed</span>
